---
date: "2021-01-01"
draft: false
excerpt: Grid is the very first CSS module created specifically to solve the layout
  problems we’ve all been hacking our way around for as long as we’ve been making
  websites.
subtitle: ""
title: Statistics - The Basics
weight: 1
---

*“There are three types of lies -- lies, damn lies, and statistics.”* ― Benjamin Disraeli

## What is Statistics?

Statistics is the process of **collecting**, **analysing** and making **inferences** from a sample of data using a parametric statistical models.  Statisticians - individuals who use statistics - attempt to draw reliable conclusions about large groups and data sets from the observable characteristics of small samples that represent only a small portion of the total population.  They do this by fitting statistical models to the sample data set and using it to estimate characteristics of the population from which the data was sampled.

## Statistical Models

A parametric statistical model for a data set consists of a **probability distribution** with unknown parameters.  Here a probability distribution is simply a mathematical function that gives the probability of occurrence of different possible outcomes of an experiment.  

## Independent and Identically Distributed

Let <img src="https://render.githubusercontent.com/render/math?math=x_1, ..., x_n"> be a data sample from a population.  A simple statistical model for this data assumes that the data can be modelled as realisations of **independent** and **identically distributed** (IID) random variables <img src="https://render.githubusercontent.com/render/math?math=X_1, ..., X_n">.

- **Independent** - 

  Knowing the random variable <img src="https://render.githubusercontent.com/render/math?math=X_i=x_i"> has no impact on the probabilities of the possible outcomes for all remaining variables.

- **Identically Distributed** - 

  Each of the random variables <img src="https://render.githubusercontent.com/render/math?math=X_1, ..., X_n"> have the same probability distribution with cumulative distribution function <img src="https://render.githubusercontent.com/render/math?math=F">.

### Example

We consider the historical daily closing price of Apple stock during the period from 01-01-2017 to 01-03-2018.  We import the time series data from Yahoo Finance and produce a plot.

```{r, include=FALSE, echo=FALSE}

## Import libraries
library(tidyquant)
library(quantmod)
library(tseries)
library(ggplot2)
library(gridExtra)

```

```{r, include=TRUE, echo=FALSE}

## Import apple stock price data
options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

# Downloading Apple price using quantmod
apple = getSymbols("AAPL", from = "2017-01-01",
           to = "2018-03-01",warnings = FALSE,
           auto.assign = FALSE)

# Define our dataset
data = as.numeric(apple$AAPL.Close)
time = 1:length(data)
plot1data = data.frame(time, data)
plot1 = ggplot(plot1data, aes(x = time, y = data)) +
  geom_line()+ggtitle("Apple stock daily closing price from 2017-01-01 to 2018-03-01.") +
  xlab("Time (days)") + ylab("Apple Stock Closing Price (Pt)")
plot1
```



```{r, include=FALSE, echo=FALSE}
logdata = ts(diff(log(data)))
time = 1:length(logdata)

plot1data = data.frame(time, logdata)

plot1 = ggplot(plot1data, aes(x = time, y = logdata)) +
  geom_line()+ggtitle("Apple stock log returns from 2017-01-01 to 2018-03-01.") +
  xlab("Time (days)") + ylab("Apple Stock Closing Price (Pt)")
plot1

plot2 = ggplot(plot1data, aes(y = logdata)) + geom_histogram(bins=30) +
  ggtitle("Histogram of log(Q)") +
  xlab("log(Q)") + ylab("Frequency")
plot2
grid.arrange(plot1, plot2, ncol=2)
```

