---
title: "Implementing Black Scholes & Vasiciek Models."
categories:
- Quantitative Analysis
- Stochastic Processes
- Financial Mathematics
author: "John Inston"
date: "2020-11-30"
excerpt: In this project I implement the Black-Scholes and Vasiciek models to determine the prices of call options and bonds respectively.  
output:   
  html_document
layout: single
header-includes:
    - \usepackage{setspace}\onehalfspacing
---

### Abstract

In this project we investigate the properties of financial stochastic processes.  We first consider the implementation of the Black-Scholes model for determining the price of European call options on stock before then utlising the Vasicek model to calculate the future price of bonds.  We discuss the background, benefits and drawbacks of both models before conducting our investigation and summarising our findings.

<br>
<br>

---

## 1.  Introduction

The Black-Scholes model is a mathematical model developed by the economists Fischer Black, Myron Scholes and Robert Merton which is widely used in the pricing of options contacts. The model assumes that the price of heavily traded assets follows a geometric Brownian motion with constant drift and volatility. When applied to a stock option, the model incorporates the constant price variation of the stock, the time value of money, the option’s strike price, and the time to the option’s expiry.
<br>

The Vasicek model is a mathematical model used in financial economics to estimate potential pathways for future interest rate changes. The model states that the movement of interest rates is affected only by random (stochastic) market movements and models interest rate movements as a factor composed of market risk, time, and equilibrium value - where the rate tends to revert towards the mean of those factors over time.
<br>

We will first investigate how the price of a European call option - as given by the Black- Scholes model - varies with changes in time, interest rates, strike price and volatility. We will then simulate an Ornstein-Uhlenbeck process and use this to simulate the price of a bond - using the Vasicek model - over a given time period before evaluating the distribution of the simulated prices.

<br>

## 2.  The Black-Scholes Model

The price at time <img src="https://render.githubusercontent.com/render/math?math=t_0=0"> of a European call option ('ECO') on a stock with strike price <img src="https://render.githubusercontent.com/render/math?math=c">, expiry time <img src="https://render.githubusercontent.com/render/math?math=t_0">, initial stock price <img src="https://render.githubusercontent.com/render/math?math=S_0">, interest rate <img src="https://render.githubusercontent.com/render/math?math=\rho"> and volatility <img src="https://render.githubusercontent.com/render/math?math=\sigma"> is given by the Black-Scholes ('BS') formula.  This is given below in Equation 1.

<br>

<img src="https://render.githubusercontent.com/render/math?math=\Large P_{t_0}=S_0\Phi\big(\frac{\log(S_0/c)%2B(\rho%2B\sigma^2/2)t_0}{\sigma\sqrt{t_0}}\big)-c\exp(\rho t_0)\Phi\big(\frac{\log(s_0/c)%2B(\rho-\sigma^2/2)t_0}{\sigma\sqrt{t_0}}\big).">

<br>

We plot the price over time of the ECO <img src="https://render.githubusercontent.com/render/math?math=P_t"> for <img src="https://render.githubusercontent.com/render/math?math=0\leq t\leq 10"> with <img src="https://render.githubusercontent.com/render/math?math=s_0=1">, <img src="https://render.githubusercontent.com/render/math?math=\sigma^2=0.02">, <img src="https://render.githubusercontent.com/render/math?math=\rho=0.03"> and <img src="https://render.githubusercontent.com/render/math?math=c=1">.  This is shown below in Figure 1.

<br>

Figure 1 - ECO price <img src="https://render.githubusercontent.com/render/math?math=P_t"> over time <img src="https://render.githubusercontent.com/render/math?math=t">.

```{r FIGURE 1, include=TRUE, echo=FALSE, fig.align='center'}
S0=1 
sigma=sqrt(0.02)
rho=0.03
c=1
t=seq(0, 10, by=0.001)

#Define our formula for the price
P=S0*pnorm((log(S0/c)+(rho+(sigma^2)/2)*t)/(sigma*sqrt(t)))-(c*exp(-rho*t))*pnorm((log(S0/c)+(rho-(sigma^2)/2)*t)/(sigma*sqrt(t)))
plot(t[-1], P[-1],  type="l",xlab="Time t", ylab="Price P at time t", main="Price P against time t")
```

<br>

We see that the price <img src="https://render.githubusercontent.com/render/math?math=P_t"> is increasing over time, which makes sense as the values of the ECO should be dependent on the time the underlying stock has to increase in value.  From Equation 1 we can see that <img src="https://render.githubusercontent.com/render/math?math=P_t\rightarrow S_0"> as <img src="https://render.githubusercontent.com/render/math?math=t\rightarrow\infty">.

We plot the price <img src="https://render.githubusercontent.com/render/math?math=P_{10}"> at time <img src="https://render.githubusercontent.com/render/math?math=t=10"> as we vary each of <img src="https://render.githubusercontent.com/render/math?math=\sigma">, <img src="https://render.githubusercontent.com/render/math?math=\rho"> and <img src="https://render.githubusercontent.com/render/math?math=c"> in turn - shown in Figure 2, Figure 3 and Figure 4 respectively.

<br>

Figure 2 - ECO price <img src="https://render.githubusercontent.com/render/math?math=P_{10}"> at time <img src="https://render.githubusercontent.com/render/math?math=t=10"> for various volatility levels <img src="https://render.githubusercontent.com/render/math?math=\sigma">.

```{r FIGURE 2, include=TRUE, echo=FALSE, fig.align='center'}

#Set t=10
t10=10

sigma1=seq(0,2, by=0.01)

P1=S0*pnorm((log(S0/c)+(rho+(sigma1^2)/2)*t10)/(sigma1*sqrt(t10)))-(c*exp(-rho*t10))*pnorm((log(S0/c)+(rho-(sigma1^2)/2)*t10)/(sigma1*sqrt(t10)))
plot(sigma1, P1, xlim=c(0, max(sigma1)),type="l", xlab="Volatility (sigma)", ylab="Price at time t=10", main="Price P10 against volatility")

```

Figure 3 - ECO price <img src="https://render.githubusercontent.com/render/math?math=P_{10}"> at time <img src="https://render.githubusercontent.com/render/math?math=t=10"> for various interest rates <img src="https://render.githubusercontent.com/render/math?math=\rho">.

```{r FIGURE 3, include=TRUE, echo=FALSE, fig.align='center'}

rho2=seq(0, 0.8, by=0.005) #Create a vector of different rho values

P2=S0*pnorm((log(S0/c)+(rho2+(sigma^2)/2)*t10)/(sigma*sqrt(t10)))-(c*exp(-rho2*t10))*pnorm((log(S0/c)+(rho2-(sigma^2)/2)*t10)/(sigma*sqrt(t10)))
plot(rho2, P2, type="l", xlab="Interest Rate (rho)", ylab="Price at time t=10", main="Price P10 against interest rate")

```

<details>
  <summary>Click to expand!</summary>
  
  ## Heading
  1. A numbered
  2. list
     * With some
     * Sub bullets
</details>
